---
title: "Lecture 4 - Mapping with R"
output: html_notebook
---

This week we will begin to explore a number of new packages that allow us to map and access mappable data in R! This will be part of a series of lectures on mapping tools as we can't hope to cover everything in one pass. First it should be noted that maps are **NOT** required as part of your final but will come into play in one or two homework assignments. 

In order to get started we need to install the requisite packages but unlike our earlier packages these will rely on things _outside_ of R in order to work, such as GDAL. Here is how I was able to install the packages on my laptop, taken from [here](https://www.karambelkar.info/2016/10/gdal-2-on-mac-with-homebrew/) using a package management tool called [`brew`](https://brew.sh/):

**YOU MUST HAVE BREW INSTALLED FOR THE FOLLOWING BLOCK TO WORK!**

```{bash}
brew unlink gdal
brew tap osgeo/osgeo4mac && brew tap --repair
brew install jasper netcdf # gdal dependencies
brew install gdal2 --with-armadillo \
--with-complete --with-libkml --with-unsupported
brew link --force gdal2
```


```{r}
install.packages("sf")
install.packages("tigris")
install.packages("tidycensus")
install.packages("viridis")
install.packages("ggplot2")
```

```{r}
library(tigris)
library(tidyverse)
library(tidycensus)
library(sf)
library(viridis) # colors and stuff
library(ggplot2)

options(tigris_class = "sf")
```

Go get a census api key [here](http://api.census.gov/data/key_signup.html)

```{r}
census_api_key("8fb6343a8789f91e7dd2ba8688d1f3018701ad95")
```
to see fips codes we use the `fips_codes` dataset
```{r}
fips_codes
```

```{r}
sf1 <- tidycensus::load_variables(2010, dataset = "acs5")
View(sf1)
```

```{r}
ma_pop <- get_decennial("tract", variables = "H0100001", year = 2010, state = "MA", geometry = T)

ma_pop %>%
  ggplot(aes(fill = value, color = value)) + 
  geom_sf() + 
#  coord_sf(crs = 26911) + 
  scale_fill_viridis(option = "magma") + 
  scale_color_viridis(option = "magma")
```

```{r}
ma <- get_acs(geography = "county", variables = "B19013_001", state = "MA")

ma

ma %>%
  mutate(NAME = gsub(" County, Massachusetts", "", NAME)) %>%
  ggplot(aes(x = estimate, y = reorder(NAME, estimate))) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point(color = "red", size = 3) +
  labs(title = "Household income by county in Mass.",
       subtitle = "2011-2015 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)")
```

```{r}
suffolk_income <- get_acs(state = "MA", county = "Suffolk", geography = "tract", 
                  variables = "B19013_001", geometry = TRUE)

plt <- suffolk_income %>%
  ggplot(aes(fill = estimate, color = estimate)) + 
  geom_sf() + 
#  coord_sf(crs = 26911) + 
  scale_fill_viridis(option = "magma") + 
  scale_color_viridis(option = "magma")

plotly::ggplotly(plt)
```

We can also plot roads and other important geographic features with the `tigris` package
```{r}
ma_schools <- school_districts(state = "MA")

ggplot(ma_schools) +
  geom_sf()

bos_roads <- roads(state = "MA", county = "Suffolk")

ggplot(bos_roads) +
  geom_sf()
```

```{r}
bos_water <- area_water(state = "MA", county = "Suffolk")

bos_nhoods <- sf::read_sf("~/Downloads/Boston_Neighborhoods.geojson")

ggplot(bos_water) +
  geom_sf()
```


```{r}
hotline <- read_csv("../data/311.csv")

plt <- ggplot(bos_nhoods) +
  geom_sf() +
  geom_point(data = hotline, aes(x = Longitude, y = Latitude, color = SUBJECT))

plotly::ggplotly(plt)
```