---
title: "BDfC - Lecture 4"
output: html_notebook
---

## String Manipulation

We've covered character vectors thus far, individual entries within a character vector are often and commonly referred to as "strings". In order to clean a lot of the data we will be working with we need to dive a bit into string manipulation.

```{r}
library(tidyverse)
library(stringr)
library(lubridate)

hotline_data <- read_csv("../data/311.csv")
biz_license <- read_csv("../data/BusinessLicenses_raw.csv")

head(hotline_data)
```

The two kinds of string manipulation we will want to do as part of this course fall into detection of strings and substituting subsections of strings.

Let's cover substitution/replacement first.
```{r}
table(hotline_data$neighborhood)

hotline_data <- hotline_data %>%
  mutate(neighborhood = str_replace(string = neighborhood, pattern = " / Brighton", "")) # removal is the same as replacing with an empty string

table(hotline_data$neighborhood)
```

Now let's cover detection, this gives us TRUE/FALSE when a string contains something that matches a pattern.

```{r}
table(biz_license$CITY)

biz_license %>%
  select(CITY) %>%
  mutate(has_hill_in_nhood = str_detect(string = CITY, pattern = "Hill")) %>%
  filter(has_hill_in_nhood)

# we can use a thing called regular expressions to make this more felxible
city_with_hill <- biz_license %>%
  select(CITY) %>%
  mutate(has_hill_in_nhood = str_detect(string = CITY, pattern = "(Hill|hill|HILL)")) %>%
  filter(has_hill_in_nhood)

table(city_with_hill$CITY)
```

## Joins and stacking

Combining data is one of the most important things you can do to be more effective in your analyses. this allows disparate data to come together and highlight things a single dataset never could.

First let's cover combining two small datasets row-wise where they have matching columns

```{r}
allston <- hotline_data %>%
  filter(neighborhood == "Allston")

allston$is_allston <- TRUE

# combines cols intelligently, even reordering. Columns that don't exists in one are added and set to NA in places where there is no data available.
bind_rows(hotline_data, allston)
```

```{r}
#joins work on the level of matching keyed values between data, ie when two datasets share a column (called a key in data terms)
licenses_by_day <- biz_license %>%
  filter(str_detect(tolower(CITY), "allston")) %>%
  mutate(neighborhood = "Allston",
         issue_date = lubridate::dmy_hm(ISSDTTM)) %>%
  group_by(day = lubridate::date(issue_date)) %>%
  count %>%
  rename(licenses_issued = n)

calls_by_day <- hotline_data %>%
  filter(str_detect(tolower(neighborhood), "allston")) %>%
  mutate(neighborhood = "Allston",
         day = lubridate::date(open_dt)) %>%
  group_by(day) %>%
  count %>%
  rename(n_calls = n)

calls_and_licenses <- inner_join(calls_by_day, licenses_by_day, by = "day")

ggplot(calls_and_licenses, aes(x = n_calls, y = licenses_issued)) +
  geom_point()
```


















