---
title: "BDfC - Lecture 4"
output: html_notebook
---

## Refresher

### loading libraries
```{r}
# only needs to be done once, has dplyr, ggplot, %>%, and readr in it!
library(tidyverse)
```

### Assignment
```{r}
n_cats <- 10
paste("I have", n_cats, "cats")
```

### Pipes
```{r}
mayors_hotline <- read_csv("../data/311.csv")

mayors_hotline %>%
  group_by(neighborhood) %>%
  count

# SAME AS

grouped_mayors <- group_by(mayors_hotline, neighborhood)
count(grouped_mayors)

mayors_hotline %>%
{print(.)} %>%
  group_by(neighborhood) %>%
  {print(.)} %>%
  count
```

```{r}
# prints but **does not assign**
mayors_hotline %>%
  group_by(neighborhood) %>%
  count

# notice the <- 
calls_by_nhood <- mayors_hotline %>%
  group_by(neighborhood) %>%
  count

calls_by_nhood  
```

### Control flow and logic

```{r}
n_cats > 10

x <- 5:15

n_cats > x
```

```{r}
if (n_cats > 10) {
  print("...You have a problem")
} else {
  print("Get more cats!")
}

ifelse(n_cats > 10, "...You have a problem", "Get more cats!")
```


### "Clean" plots

Clean plots are very important and you will be graded as such! If noone can read your plot, it's not going to be very helpful. Let's look at some ways to fix.

```{r}
calls_by_nhood

ggplot(data = calls_by_nhood, mapping = aes(x = neighborhood, y = n)) +
  geom_bar(stat = "identity")
```

looks messy! Some problems: no title, can't read x axis, bad x/y labels. Nice to have: reorder cols. 

```{r}
# one way to deal with x axis overlap, if it's not too severe
ggplot(data = calls_by_nhood, mapping = aes(x = neighborhood, y = n)) +
  geom_bar(stat = "identity") +
  # flip x and y
  coord_flip()
```

Anoter way to fix is by rotating the labels

```{r}
ggplot(data = calls_by_nhood, mapping = aes(x = neighborhood, y = n)) +
  geom_bar(stat = "identity") +
  # I personally have to look up this syntax every single time
  theme(axis.text.x = element_text(angle = 65, hjust = 1))
```

adding/fixing labels is simple, too.

```{r}
# one way to deal with x axis overlap, if it's not too severe
ggplot(data = calls_by_nhood, mapping = aes(x = neighborhood, y = n)) +
  geom_bar(stat = "identity") +
  # flip x and y
  coord_flip() +
  # x label
  xlab("") +
  ylab("Count of Cases Created")
```

Same with titles!
```{r}
ggplot(data = calls_by_nhood, mapping = aes(x = neighborhood, y = n)) +
  geom_bar(stat = "identity") +
  # flip x and y
  coord_flip() +
  # x label
  xlab("") +
  ylab("Count of Cases Created") +
  ggtitle("Count of 311 Cases Created by Neighborhood")
```

Reordering has kind of a wacky syntax (arrange does not work!), but it's simple once you see it.

```{r}
ggplot(data = calls_by_nhood, mapping = aes(x = reorder(neighborhood, n), y = n)) +
  geom_bar(stat = "identity") +
  # flip x and y
  coord_flip() +
  # x label
  xlab("") +
  ylab("Count of Cases Created") +
  ggtitle("Count of 311 Cases Created by Neighborhood")
```

We can add themes to further style:
```{r}
base_plot <- ggplot(data = calls_by_nhood, mapping = aes(x = reorder(neighborhood, n), y = n)) +
  geom_point() +
  # flip x and y
  coord_flip() +
  # x label
  xlab("") +
  ylab("Count of Cases Created") +
  ggtitle("Count of 311 Cases Created by Neighborhood")

base_plot

base_plot +
  theme_bw()

base_plot +
  theme_classic()

base_plot +
  theme_minimal()

clean_plot <- base_plot +
# taken from https://ase.tufts.edu/bugs/guide/assets/R%20Graphics%20Cookbook.pdf USE THIS BOOK!
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

library(plotly)

ggplotly(clean_plot)
```


## Vectors

Vectors in R are the base of everything we will work with. Vectors are containers for one or more than piece of data of the type NA, chracter, numeric, or factor.

We can make vectors with `c()`
```{r}
c("this", "is", "a", "character", "vector")
c(1:10)

x <- 1

is.vector(x)
```

Vectors have a notion of length and "indicies" or numbers that enumerate upwards for each "slot" in a vector. In R our indices start with 1.

```{r}
x <- c(10:25)

length(x)

# this is called subsetting a vector
x[1]


x[c(1,10)]
```

We subset by either supplying numeric indicies or logical vectors

```{r}
library(tidyverse)

x[15]

x[x >= 20]

service_requests <- read_csv("../data/311.csv")

service_requests$open_dt[service_requests$open_dt > lubridate::ymd("2017-09-01")]
```

These indicies and a notion of distinct slots for each piece of data in a container allows us to begin talking about iteration or a means to iteratively apply an operation accross a container like a vector or even an entire data frame.

# NOT GOING OVER, NO NEED TO READ ITERATION CHAPTER

## Iteration

Iteration allows us to iteratively apply functions accross containers such as vectors or dataframes. You have seen some of this already with `group_by` amd `summarise` where functions are applied accross each group. Iteration helps with the DRY principle we talked about last week. By putting our repeated logic in functions and iterating across our containers applying those functions when a condition is met allows us to reduce a large amount of duplicate code.
```{r}
# turn all character data to lowercase only
service_requests %>%
  map_if(is.character, tolower) %>%
  data.frame()
```

map mostly helps us with applying functions over collections of data frames. For this we will need a list of data frames

```{r}
nhood_list <- service_requests %>%
  split(.$neighborhood)

nhood_list
```


```{r}
service_requests %>%
  split(.$neighborhood) %>%
  map(summary)
```


```{r}
geo_plot  <- function(df, lat, long, color_by, plot_dir, save_plot=TRUE) {
  # get nhood 
  nhood = unique(df$neighborhood)
  
  plt = ggplot(df, aes_string(x = long, y = lat, color = color_by)) +
    geom_point() +
    ggtitle(nhood)
  return(plt)
}


paste0("this", "is", )
table(service_requests$Department)

service_requests %>%
  split(.$neighborhood) %>%
  map(filter, CASE_STATUS == "Closed") %>%
  map(geo_plot, "Latitude", "Longitude", "SUBJECT", "../plots/")

service_requests %>%
  split(.$neighborhood) %>%
  map(filter, CASE_STATUS == "Closed") %>% 
  map(filter, Department == "BWSC") %>%
  map(geo_plot, "Latitude", "Longitude", "REASON")

# save to disk
geo_plot  <- function(df, lat, long, color_by, plot_dir, save_plot=TRUE) {
  # get nhood 
  nhood = unique(df$neighborhood)
  
  plt = ggplot(df, aes_string(x = long, y = lat, color = color_by)) +
    geom_point() +
    ggtitle(nhood)
  if (save_plot) {
    ggsave(plot = plt, filename = paste0(plot_dir, nhood, ".jpg", collapse = ''), width = 9)
  }
  return(plt)
}

```



